---
title             : " Supplement to 'Common or Distinct Attention Mechanisms for Contrast and Assimilation?'"
shorttitle        : "Contrast and Assimilation Supplement"

author: 
  - name          : "Hope K. Snyder"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : " "
    email         : "hks7w2@mail.missouri.edu"
  - name          : "Sean M. Rafferty"
    affiliation   : "1"
  - name          : "Julia M. Haaf"
    affiliation   : "1"
  - name          : "Jeffery N. Rouder"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "University of Missouri"

author_note: >
  This document was written in R-Markdown with code for data analysis integrated into the text.  The Markdown script is open and freely available at https://github.com/PerceptionAndCognitionLab/ctx-flanker/tree/public/papers/current.  The data were *born open* [@Rouder:2016] and are freely available at https://github.com/PerceptionCognitionLab/data1/tree/master/ctxIndDif/flankerMorph4

keywords          : "Inhibition, Selective Attention, Contrast Effects, Assimilation Effects"
wordcount         : " "

bibliography      : ["r-references.bib","bibfile.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no

lang              : "english"
class             : "man"
header-includes:
   - \usepackage{enumerate}
   - \usepackage{amsmath, amssymb,amsthm,amsfonts}
   - \usepackage{bm}
   - \usepackage{multicol}
   - \usepackage[usenames,dvipsnames]{pstricks}
   - \usepackage{mathtools}
output            : papaja::apa6_pdf
---

```{r include = FALSE}
library("papaja")
```

This document is the supplement to "Common or Distinct Attention Mechanisms for Contrast and Assimilation?".   It provides the specification and analysis of a hierarchical Bayesian probit model for assessing the correlation across individuals' abilities to inhibit distractors in assimilation and contrast contexts.

# Model Specification

Let $Y=0,1$ denote whether a response is "A" or "H", respectively.  Further, let $Y_{ijk\ell m}$ denote the response for the  $i$th participant, $i=1,\ldots,I$, in the $j$th context type ($j=1,2$ for word and letter contexts, respectively), for the $k$th context direction ($k=1,2$, for contexts that promote "A"  and "H" responses, respectively), for the  $\ell$th target, $\ell=1,\ldots,L$, and for the $m$th replicate, $m=1,\ldots ,M_{ijk\ell}$.  The context assignments are displayed in the following table:

```{r table,echo = FALSE,cache=TRUE}
library(pander)
quest6 <- matrix(c(NA,"A-promoting","H-promoting", "word","C_T","T_E","letter","H","A"),ncol=3,byrow=TRUE)
colnames(quest6) <- c(" ","k=1","k=2")
rownames(quest6) <- c(" ","j=1","j=2")
quest6 <- as.table(quest6)

panderOptions("missing", " ")
pander(quest6)
```

Observations $Y_{ijk\ell m}$ are dichotomous.  To model the effect of covariates in dichotomous observations, we use a probit-regression specification:
\[Y_{ijk\ell m} \stackrel{ind}{\sim} \text{Bernoulli}\left[\Phi(\mu_{ijk\ell})\right].\]

Here, $\Phi$, the cumulative distribution function of the standard normal, is the link, and $\mu_{ijk\ell} \in (\infty,\infty)$ is the combined effect of people, conditions, and the target. 

To model individual inhibition effects, we additively decompose $\mu_{ijk\ell}$ into the effect of the target for a particular participant, $\gamma_{i\ell}$, the individual's *assimilation* effect when the background context is a word, $\alpha_i$ and the individual's *contrast* effect when the background context is a letter frame, $\beta_i$.  The decomposition is:
\[\mu_{ijk\ell}=\gamma_{i\ell}+v_jx_k\alpha_i+(1-v_j)x_k\beta_i.\]


The quantities $v_j$ and $x_k$ are indicators of the context type and direction, as follows:
\[ v_j=
  \begin{dcases}
      0 & \quad \text{if} \quad j =2  \text{ (letters)}\\
      1 & \quad \text{if} \quad j =1 \text{ (word)}
  \end{dcases}
  \qquad
  x_k=
  \begin{dcases}
      \frac{-1}{2} & \quad \text{if} \quad k =1  \text{ (A-promoting)}\\
      \frac{1}{2}  & \quad \text{if} \quad k =2 \text{ (H-promoting)}
  \end{dcases}
\]

Prior distributions are needed for $\alpha_i, \beta_i, \text{ and } \gamma_{i\ell}$.  We start with $\gamma_{i\ell}$, the effect of the target for a particular participant.  These parameters are not of primary concern, and we use a broad hierarchical prior given by: $\gamma_{i\ell} \sim \mbox{N}(\nu_\ell,\delta_\ell)$ with hyper priors of $\nu_{\ell} \sim \mbox{N}(0,1)$ and $\delta_{\ell} \sim \mbox{Inverse Gamma}(.5,.01)$.  
The effects of interest are an individual's assimilation effect, $\alpha_i$, and an individual's contrast effect, $\beta_i$.  We follow the development in @Rouder:etal:2007a who describe Bayesian analyses in these settings.  The joint prior over these parameter is
\[\begin{pmatrix}
  \alpha_i\\
  \beta_i
\end{pmatrix} = \mbox{N}(\boldsymbol{\lambda},\boldsymbol{\Sigma}),\] 
where $\boldsymbol{\lambda}$ is the vector $(\mu_{\alpha},\mu_{\beta})$ and $\boldsymbol{\Sigma}=\begin{pmatrix} \Sigma_{1,1} & \Sigma_{1,2} \\ \Sigma_{2,1} &  \Sigma_{2,2} \end{pmatrix}$ is a covariance matrix.

Following @Rouder:etal:2007a, the hyperpriors for $\mu_{\alpha}$ and $\mu_{\beta}$ are given by
\[\mu_{\alpha} \sim N(0,1),\]
\[ \mu_{\beta} \sim N(0,1). \]
The hyperprior for $\boldsymbol{\Sigma}$ is
\[\boldsymbol{\Sigma} \sim \mbox{Inverse Wishert}(3,\Omega),\]
where $\Omega = \begin{vmatrix}.05 & 0 \\ 0 & .05 \end{vmatrix}$.
A key property of the above specification is that population covariance in $\boldsymbol{\Sigma}$ is an explicit parameter given by $\Sigma_{1,2}$.  Hence, a population correlation is well defined, and subsequent inferences generalize to new data from new participants.

The choices we make above are all reasonable given the expected degree of variability in proportions.  The standard normal prior in the probit space on $\mu_{\beta}$ and $\mu_{\alpha}$ corresponds to flat priors in the space of proportions, and the shape value of $3$ in the Inverse Wishert corresponds to a flat prior on the population-level correlation coefficient.  The only substantive choice are the values $.05$ in the scale of the Inverse Wishert.  This small value of precision corresponds to a vaguely informative prior on effects.  Additional details are provided in @Rouder:etal:2007a.

# Analyses and Results

The estimation of probit-transformed parameters follows from @Albert:Chib:1995.  Accordingly, a new set of latent variables is introduced for each observation, and we denote them $w_{ijk\ell m}$.  These variables are defined as follows:
\[y_{ijk\ell m}=
  \begin{dcases}
      0 & \text{ then }  w_{ijk\ell m} < 0 \\ 
      1  & \text{ then }  w_{ijk\ell m} \geq 0
  \end{dcases}
\]

Without loss of generality, the variable $w_{ijk\ell m}$ is assumed to be distributed as a normal with a variance of one and the mean of $\mu_{ijk\ell}$.  Thus, the parameters of interest at this level are $w_{ijk\ell m}$ and $\mu_{ijk\ell}$.  It is easy to compute the posterior of $\boldsymbol{w}$ given the vector of $\boldsymbol{\mu}$ and, conversely, it is easy to compute the posterior of $\boldsymbol{\mu}$ given the $\boldsymbol{w}$'s.  Hence, the marginals of each may be found by Markov chain Monte Carlo  (MCMC) sampling.   The conditional posterior distribution of $\boldsymbol{w}$ is a truncated normal:
\[w_{ijk\ell m}|y_{ijk\ell m} \sim 
  \begin{dcases} 
    \text{N}_-(\mu_{ijk\ell},1) & \text{ if }  y_{ijk\ell m}=0 \\
    \text{N}_+(\mu_{ijk\ell},1) & \text{ if }  y_{ijk\ell m}=1, 
  \end{dcases}
\]
where $N_+$ and $N_-$ denote normals distributions truncated at zero from below and above, respectively.

Posterior distributions for remaining parameters may also be sampled with MCMC.  In all cases, conditional posterior distributions of parameters may be derived directly from the proportional form of Bayes rule [@Jackman:2009;@Rouder:Lu:2005].  Priors were chosen to be conjugate, and consequently, posterior distributions may be sampled from known distributions in Gibbs steps. 

Critical parameters are individual estimates of contrast and assimilation, and the population-level correlation of these abilities.  Individual estimates are parameters $\alpha_i$ and $\beta_i$.  Posterior means and posterior standard deviations of these parameters are shown as points and ellipses, respectively, in Figure 4A in the main paper.  The most critical parameter is the population correlation.  On each iteration of the MCMC chain, we calculated $\rho=\Sigma_{1,2}/\sqrt{\Sigma_{1,1}\times \Sigma_{2,2}}$.  The prior and posterior distribution of $\rho$ is shown in Figure 4B.  A full description of the results may be found in the main paper.

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
